To be FAIR: Theory Specification Needs an Update

### 2025-06-08 action_letter.Rmd
update 2025-06-08 14:18:50.386429

<!-- AB: -->
3.	Regarding missing reference to "research waste" - yes, this should be picked up. By the way, the latest revision of the Declaration of Helsinki states that ethical science has to avoid "research waste". Pretty cool.

@@

```{asis}
Line 73 onwards: The example of SDT nicely illustrates the authors’ point here!
```

@@

<!-- AB: -->
4.	What is a theory? (Suggestion to move up the section) It seems to always come back to this that readers a confused because everyone has their own idea of what a theory is. We need to think of one introductory sentence very early on that basically sets the stage and tells people no matter what they think a theory is they very likely will find the FAIRification useful

@@

We removed the "cf.", although we couldn't find mention of it in the APA7 style guide. We updated the citation scheme to APA7, so three or more authors will be shortened to "et al.".

@@

We respectfully disagree; open data - one example of modular publishing - has increased dramatically in recent decades, up from ~7% of articles in 2000–2009 to ~50% of articles in 2010-2019.
Given this, we don't see a reason to note this as a limitation, but we have added a reference that indicates that data sharing has become widely adopted in psychology:

@@

@@

>>>>

@@

As requested, we have added a few more words - but we also want to avoid redundancy, as the FAIR principles are fairly (pun intended) common sense, and well-documented in other publications which are all cited in the opening paragraph. In the body of the text, we explicitly engage with each of the FAIR (sub-) principles, so all information is already contained within the paper.

To address the comment, we have rewritten the opening paragraph as follows:

`r get_revision("whatsfair")`

@@

We have done as requested, and additionally, include a sentence in the first paragraph to pre-emptively address readers' concerns that FAIR theory may only be relevant for those who have adopted a specific paradigm:

`r get_revision("alldefinitions")`

@@

We removed "cf." here, although we couldn't find mention of it in the APA7 style guide.
We use "cf." to indicate when a specific reference *contradicts* a point; e.g. "all theories should be formal (Smaldino, 2017; cf. Oude-Maatman, 2024)": Oude-Maatman disagrees with this statement.
We have also updated the citation scheme to APA7, so three or more authors will be shortened to "et al.".
Note that some journals prescribe citation styles that (partly) conflict with APA7; we defer to the Editor on how to handle such cases.

@@

We respectfully disagree; open data - one example of modular publishing - has increased dramatically in recent decades, up from ~7% of articles in 2000–2009 to ~50% of psychological articles in 2010-2019, and more (~80%) in other disciplines.
Given this, we don't see a reason to note this as a limitation, but we have added a reference that indicates that data sharing has become widely adopted in psychology:

@@

We have rewritten this section slightly, including a brief explanation of the "open empirical cycle" and how the present work relates to it:

`r get_revision("openec")`

@@

We had mentioned the learning curve in the Limitations section, and have expanded this discussion to address the present comment:

`r get_revision("learningcurve")`
